---
title: "STOIK dataset"
output: 
  html_document: 
    toc: true
    fig_height: 13
    fig_width: 14
    
---
    


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)


library(energy)

load("TOIK-data-for-maria.rdata")
library(zCompositions)
source("modifiedCmultRepl.R")
#debugged to allow using gbm prior see diff with cmultRepl

library(gtools)
library(compositions)
library(robCompositions)


```


## Bayesian-Multiplicative replacement of count zeros

Zero counts, assumed to be a consequence of the sampling process (insufficient sampling),are imputed under a Bayesian paradigm ('GBM', 'SQ' or 'BL' method)
     by posterior estimates of the multinomial probabilities generating
     the counts, assuming a Dirichlet prior distribution. The argument
     'method' sets the Dirichlet hyper-parameters 't' (priori estimates
     of multinomial probabilities) and 's' (_strength_). The user can
     specify their own by setting 'method="user"' and entering them as
     't' and 's' arguments. Note that, under certain circumstances (see
     references for details), these methods can generate imputed
     proportions falling above the lowest estimated probability of a
     multinomial part (_c_/_n_, where _c_ is the count and _n_ is the
     number of trials). In such cases, the imputation is corrected by
     using a fraction ('delta') of the minimum _c_/_n_ for that part.
     Lastly, the non-zero parts are multiplicatively adjusted according
     to their compositional nature.

     On the other hand, 'method="CZM"' uses multiplicative simple
     replacement ('multRepl') on the matrix of estimated probabilities.
     The upper limit and the fraction delta used are specified by,
     respectively, the arguments 'threshold' and 'delta'. Suggested
     values are 'threshold=0.5' (so the upper limit for a multinomial
     probability turns out to be 0.5/_n_), and 'delta=0.65' (so the


```{r whereAreWeInFileSystem}
#
getwd()

stolice_L7_abs[1:3,1:3]
```


```{r ZeroReplace}
#


 #input dataset 
BayesianMultiplicative2ReplacedCountZerosL7.r =  stolice_L7_abs 


# using GBM prior ... use the fixed version!!! 
#modifiedC
BayesianMultiplicative2ReplacedCountZerosL7.gbm <- modifiedCmultRepl(t(BayesianMultiplicative2ReplacedCountZerosL7.r),  label=0, method="GBM",output = "counts")



BayesianMultiplicative2ReplacedCountZerosL7.gbm[1:3,1:3]


# using SQ prior ...

BayesianMultiplicative2ReplacedCountZerosL7.sq <- cmultRepl(t(BayesianMultiplicative2ReplacedCountZerosL7.r),  label=0, method="SQ",output = "counts")

# using BL prior ...

BayesianMultiplicative2ReplacedCountZerosL7.bl <- cmultRepl(t(BayesianMultiplicative2ReplacedCountZerosL7.r),  label=0, method="BL",output = "counts")


# using CZM approach ...
BayesianMultiplicative2ReplacedCountZerosL7.czm <- cmultRepl(t(BayesianMultiplicative2ReplacedCountZerosL7.r),  label=0, method="CZM",output = "counts")


save(BayesianMultiplicative2ReplacedCountZerosL7.czm,BayesianMultiplicative2ReplacedCountZerosL7.bl,BayesianMultiplicative2ReplacedCountZerosL7.sq, BayesianMultiplicative2ReplacedCountZerosL7.gbm,file="ZerosReplacedinputDatasetL7.RData")

#load("ZerosReplacedinputDatasetL7.RData")

#load("ZerosReplacedinputDatasetL7.RData")

```

## Including Plots

You can also embed plots, for example:

```{r pcaANDdendro, echo=TRUE, include=TRUE}
old.par <- par()
###### create a complex object with all combinations of zeros replacement and log transformation approaches

combiObj=list()
vectorOfComb=vector()#vector(mode="numeric",length=12)
for(stringMeth in c("sq","bl","gbm","czm")){ #,"czm","clrsq","clrbl","clrgbm","clrczm"

temp.pattern=paste("*",stringMeth,"$",sep="")
string2eval=ls(pattern=glob2rx(temp.pattern))
#example of how string2eval looks like: "BayesianMultiplicativeReplaceCountZerosL7.czm"

combiObj[[stringMeth]]<-as.data.frame(eval(parse(text=string2eval)))#ls()[i]

cat(stringMeth,"\n")
show(dim(combiObj[[stringMeth]]))

## applying total sum scaling approach : tss values wanted to have compositions

#combiObj[[stringMeth]][[stringMeth]] <- combiObj[[stringMeth]]

# this does not work because apply does some unwanted extra:  combiObj[[stringMeth]][["test"]] <- apply(combiObj[[stringMeth]][[stringMeth]], 1, function(x){x/sum(x)})

# this works fine

tempLabelForProp=paste(stringMeth,'tss',sep="_")

#combiObj[[stringMeth]][[tempLabelForProp]]<-combiObj[[stringMeth]][[stringMeth]]/rowSums(combiObj[[stringMeth]][[stringMeth]])

combiObj[[tempLabelForProp]]<-combiObj[[stringMeth]]/rowSums(combiObj[[stringMeth]])

#show(dim(combiObj[[stringMeth]][[stringMeth]]))
show(dim(combiObj[[stringMeth]]))
show(dim(combiObj[[tempLabelForProp]]))


vectorOfComb=c(vectorOfComb,tempLabelForProp)


## log transform
# applying log transform to compositions ( total sum scaling values ) 
# the next steps...for each zeros replacement method you want to apply a log transform approach...

for(logTransform in c("cenLR","addLR")){#,"clr","alr") 

string2eval2=paste("combiObj[[\"",tempLabelForProp,"\"]]",sep="")
#string2eval2
x<-eval(parse(text =paste(logTransform,"(",string2eval2,")")))

#combiObj[[tempLabelForProp]][["LOGT"]]=list()
tempLabelForPropLogT=paste(tempLabelForProp,logTransform,sep="_")
combiObj[[tempLabelForPropLogT]] <- as.data.frame(x$x) #eval(parse(text =paste(logTransform,"(",string2eval,")"))))

vectorOfComb=c(vectorOfComb,tempLabelForPropLogT) # paste(stringMeth,logTransform,sep="_")

}

for(logTransform in c("ilr")){#,"clr","alr")

string2eval2=paste("combiObj[[\"",tempLabelForProp,"\"]]",sep="")

tempLabelForPropLogT=paste(tempLabelForProp,logTransform,sep="_")
combiObj[[tempLabelForPropLogT]] <- eval(parse(text =paste(logTransform,"(",string2eval,")")))
 
vectorOfComb=c(vectorOfComb,tempLabelForPropLogT) #paste(stringMeth,logTransform,sep="_")
 
}

}


vectorOfComb

### create different kind of inputs for multivariate analysis and comparison....both isolation kits .....
### all the 3 different sample collection strategies

################ visualization by pca
#### hierarchical clustering

EHCres=list()
mat<- matrix(NaN, nrow = 16, ncol=2, byrow = TRUE, dimnames = list(seq(1:16) ,c("PC1","PC2")))
rownames(mat)<-vectorOfComb

for(combi in vectorOfComb ){

subsetIdx.1=grep("\\.S1\\.1$",rownames(combiObj[[combi]]))

subsetIdx.2=grep("\\.S2\\.1$",rownames(combiObj[[combi]]))

subsetIdx.3=grep("\\.S3\\.1$",rownames(combiObj[[combi]]))

subsetIdx.4=grep("\\.S1\\.2$",rownames(combiObj[[combi]]))

subsetIdx.5=grep("\\.S2\\.2$",rownames(combiObj[[combi]]))

subsetIdx.6=grep("\\.S3\\.2$",rownames(combiObj[[combi]]))


df2prcomp=rbind(

combiObj[[combi]][subsetIdx.1,],
combiObj[[combi]][subsetIdx.2,],
combiObj[[combi]][subsetIdx.3,],
combiObj[[combi]][subsetIdx.4,],
combiObj[[combi]][subsetIdx.5,],
combiObj[[combi]][subsetIdx.6,]

)

subsetIdx=c(subsetIdx.1,subsetIdx.2,subsetIdx.3,subsetIdx.4,subsetIdx.5,subsetIdx.6)

pcx.abund <- prcomp(df2prcomp)

mat[rownames(mat)==paste(stringMeth,logTransform,sep="_"),colnames(mat)=="PC1"]<-pcx.abund$sdev[1]^2/mvar(combiObj[[combi]])
mat[rownames(mat)==paste(stringMeth,logTransform,sep="_"),colnames(mat)=="PC2"]<-pcx.abund$sdev[2]^2/mvar(combiObj[[combi]])

palette=colors()[2:17]  #palette(c( rgb(0,0,1,0.6), rgb(1,0,0,0.6),rgb(.3,0,.3,0.6)))

layout(matrix(c(1,2),1,2, byrow=T), widths=c(6,2), heights=c(8,3))

par(mgp=c(2,0.5,0))

plot(pcx.abund$x,col=c("yellow4","red","pink","brown","magenta","green","blue","grey70","black","coral","lavender","red4","gold","purple","olivedrab","turquoise2")[as.numeric(as.factor(substr(rownames(combiObj[[combi]][subsetIdx,]), 2, 4) ))],pch=c(21, 18, 24, 20,23,25)[as.numeric(as.factor(substr(rownames(combiObj[[combi]][subsetIdx,]), 7, 9)))],#as.vector(conds$cond)


xlab=paste("PC1 ", round (sum(pcx.abund$sdev[1]^2)/mvar(combiObj[[combi]]),3), sep=""),


ylab=paste("PC2 ", round (sum(pcx.abund$sdev[2]^2)/mvar(combiObj[[combi]]),3), sep=""),


main=paste("Biplot data transf. -", combi, sep=""))


legend("topleft", pch =c(21,24,23,20,25,18) , col="black", legend = unique((as.factor(substr(rownames(combiObj[[combi]][subsetIdx,]), 7, 9)))))

#text(pcx.abund$x[,1],pcx.abund$x[,2],labels=substr(rownames(pcx.abund$x), 2, 4))

barplot(pcx.abund$sdev^2/mvar(combiObj[[combi]]), ylab="variance explained",xlab="Component", main="Scree plot") # scree plot

fn=paste("L7_",combi, ".pdf" ,sep="") #  stringMeth, "_" , logTransform
dev.copy(pdf,file=fn)
dev.off()

EHCres[[combi]][[1]]<-energy.hclust(dist(combiObj[[combi]][subsetIdx,]))

# the ground truth : 16 individuals each associated with 3 different sample collection method and 2 isolation kits; 16*3*2=96 

g <- setNames(c(1:96), rownames(combiObj[[combi]])[subsetIdx]) 
for(i in 1:16){
#cat(i,"\n")
togrep=substr(names(cutree(EHCres[[combi]][[1]], k=16))[i],2,4)
mysubset=grep(togrep,names(cutree(EHCres[[combi]][[1]], k=16)))
g[mysubset]<-i
}

par(cex=0.7, mar=c(15, 8, 4, 1))
plot(EHCres[[combi]][[1]],hang=-1)
EHCres[[combi]][[2]]<-table(cutree(EHCres[[combi]][[1]], k=16) == g)
cat(combi,"\n")
show(EHCres[[combi]][[2]])
}



```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
